### 2023-06-23 Draft.tex
Written separable filtering

>>>>

Also notated as $K=vh^{\pmb{T}}$, where $\pmb{T}$ denotes tranposing $h$.

A method to check whether a given Kernel is separable is to use \emph{Singular Value Decomposition (SVD)}\cite{Szeliski_2022}

\[        
    \pmb{K} = \sum_{i}{\sigma_i \cdot u_i \cdot v_i^{\pmb{T}}} 
\]
or more generally
\begin{align*}
        \pmb{K} =& \textbf{U}_{m\times p} \pmb{\textstyle\sum}_{p\times p} \textbf{V}_{p\times n}^{\pmb{T}} \\
        \pmb{K} =& \begin{bmatrix}
            u_0 & \hdots & u_{p-1}
        \end{bmatrix}
        \begin{bmatrix}
            \sigma_0&&\\
            &\ddots&\\
            &&\sigma_{p-1}
        \end{bmatrix}
        \begin{bmatrix}
            v_0 \\
            \vdots \\
            v_{p-1}
        \end{bmatrix}
\end{align*}
Note that it requires the number of rows and columns must be the same, we can also alleviate this by putting 0 appropriately,
to allow the kernel to be a square matrix.Then the first singular value $\sigma_0$ is checked. If it is the only one that is non zero,
and the rest is zero, then the kernel is separable. As an example, the sobel filter, which is used for edge detection, 
can be separated into two separate operations. In this case, this is for the horizontal derivative approximation.
\[
    \frac{1}{8}
    \begin{bmatrix}
        -1 & 0 & 1 \\
        -2 & 0 & 2 \\
        -1 & 0 & 1
    \end{bmatrix}
    =
    \frac{1}{8}
    \begin{bmatrix} -1 & 0 & 1 \end{bmatrix}
    \begin{bmatrix} 1 \\ 2 \\ 1 \end{bmatrix}
\]
The normalization of the operators are easily calculated by dividing each operator with the absolute sum of the entries.
\[  
    \begin{bmatrix} -1 & 0 & 1 \end{bmatrix} \Rightarrow \vert -1 \vert + \vert 0 \vert + \vert 1 \vert = 2 \Rightarrow \frac{1}{2}
\]
Then the normalized operator is
\[
    \frac{1}{2} \begin{bmatrix} -1 & 0 & 1 \end{bmatrix}
\]
it applies to the vertical operator as well
\[
    \frac{1}{4} \begin{bmatrix} 1 \\ 2 \\ 1 \end{bmatrix}
\]
The following are more examples of common separable filtering.

\begin{center}\begin{tabular}{c c c c c}
    $
        \frac{1}{K^2}
        \begin{bmatrix}
           1 & 1 & \hdots & 1 \\
           1 & 1 & \hdots & 1 \\
           \vdots & \vdots & \ddots & \vdots \\
           1 & 1 & \hdots & 1
        \end{bmatrix}
    $
 &
    $
        \frac{1}{16}
        \begin{bmatrix}
            1 & 2 & 1 \\
            2 & 4 & 2 \\
            1 & 2 & 1
        \end{bmatrix}
    $
& 
    $
        \frac{1}{256}
        \begin{bmatrix}
            1 & 4 & 6 & 4 & 1 \\
            4 & 16 & 24 & 16  & 4 \\
            6 & 24 & 36 & 24 & 6 \\
            4 & 16 & 24 & 16  & 4 \\
            1 & 4 & 6 & 4 & 1
        \end{bmatrix}
    $
& 
    $
        \frac{1}{8}
        \begin{bmatrix}
            -1 & 0 & 1 \\
            -2 & 0 & 2 \\
            -1 & 0 & 1
        \end{bmatrix}
    $
& 
    $
        \frac{1}{4}
        \begin{bmatrix}
            1 & -2 & 1 \\
            -2 & 4 & -2  \\
            1 & -2 & 1
        \end{bmatrix}
    $
\\
    $
        \frac{1}{K} 
        \begin{bmatrix} 1 & 1 & \hdots & 1 \end{bmatrix}
    $
&
    $
        \frac{1}{4} 
        \begin{bmatrix} 1 & 2 & 1 \end{bmatrix}
    $
&   
    $
        \frac{1}{16}
        \begin{bmatrix} 1 & 4 & 6 & 4 & 1 \end{bmatrix}
    $
&   
    $
        \frac{1}{2} 
        \begin{bmatrix} -1 & 0 & 1 \end{bmatrix}
    $
&
    $
        \frac{1}{2} 
        \begin{bmatrix} -1 & 2 & 1 \end{bmatrix}
    $
\\
Box & Bilinear & Gaussian & Sobel & Corner 
\end{tabular}\end{center}

Note that for Corner filtering, the kernel is not normalized.
With some filters, it is not always possible to separate it into just two operations, therefore it can be
transformed into a series of 1 dimensional operations. However, this may be impractical as it increases the time and space
complexity.
